---
title: "Lab 10"
author: "ks"
date: "11/05/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```

## Lab week 10 

```{r install-libraries}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)
```

```{r create-temp-database}
# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")
```


```{r tables}
# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")
```


```{r database}
# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)
```

```{r check}
dbListTables(con)
```

You can also use knitr + SQL!

TIP: Use can use the following QUERY to see the structure of a table.

```{sql, connection=con,output.var="x1"}
PRAGMA table_info(actor)
```

```{r}
x1
```
This is equivalent to using `dbGetQuery`.

```{r}
dbGetQuery(con, "PRAGMA table_info(actor)")
```


SQL references:  
https://www.w3schools.com/sql/

## Exercise 1  
Retrive the actor ID, first name and last name for all actors using the actor table. Sort by last name and then by first name.

Using LIMIT n, we can print just the first n entries.

```{r ex1}
dbGetQuery(con, " 
SELECT actor_id, first_name, last_name
FROM actor
ORDER by last_name, first_name
LIMIT 6")
```



## Exercise 2
Retrive the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.

```{r Ex2.1}
dbGetQuery(con, " 
SELECT actor_id, first_name, last_name
FROM actor
WHERE last_name IN ('WILLIAMS', 'DAVIS')
")
```

This also works:
```{r Ex2.2}
dbGetQuery(con, " 
SELECT actor_id, first_name, last_name
FROM actor
WHERE last_name='WILLIAMS' OR last_name='DAVIS'
")
```


## Exercise 3
Write a query against the rental table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID.

```{r}
dbGetQuery(con, "PRAGMA table_info(rental)")
```


```{r Ex3}
dbGetQuery(con, " 
SELECT DISTINCT customer_id 
FROM  rental
WHERE date(rental_date) = '2005-07-05'
LIMIT 6")
```


## Exercise 4
### Exercise 4.1
Construct a query that retrieves all rows from the payment table where the amount is either 1.99, 7.99, 9.99.

```{r}
dbGetQuery(con, "PRAGMA table_info(payment)")
```

If you have a really big dataset you may want to do it this way:

```{r Ex4.1}
qq<- dbSendQuery(con, " 
SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)
")
dbFetch(qq,n=10)
```
Now we can grab the next 10 and then close.
```{r next10}
dbFetch(qq,n=10)
dbClearResult(qq)
```

### Exercise 4.2
Construct a query that retrieves all rows from the payment table where the amount is greater than 5.


```{r Ex4.2}
dbGetQuery(con, " 
SELECT *
FROM payment
WHERE amount > 5
LIMIT 10
")
```
```{r Ex4.2.2}
dbGetQuery(con, " 
SELECT staff_id, COUNT(*)
FROM payment
/* GROUP BY goes AFTER WHERE*/
WHERE amount > 5
GROUP BY staff_id
")
```

### Exercise 4.3
Construct a query that retrieves all rows from the payment table where the amount is greater then 5 and less then 8.


```{r Ex4.3}
dbGetQuery(con, " 
SELECT *
FROM payment
WHERE amount > 5 AND amount < 8
LIMIT 6")
```



## Exercise 5
Retrieve all the payment IDs and their amount from the customers whose last name is ‘DAVIS’.


```{r Ex5}
dbGetQuery(con, " 
SELECT last_name, payment_id, amount 
FROM customer AS a INNER JOIN payment AS b 
  ON a.customer_id = b.customer_id
WHERE last_name IS 'DAVIS'
")
```


## Exercise 6
### Exercise 6.1
Use COUNT(*) to count the number of rows in rental.
```{r ex6.1}
dbGetQuery(con,"
 SELECT COUNT(*)
 FROM rental
")
```


### Exercise 6.2
Use COUNT(*) and GROUP BY to count the number of rentals for each customer_id.
```{r ex6.2}
dbGetQuery(con,"
 SELECT customer_id, COUNT(*) AS  `N Rentals`
 FROM rental
 GROUP BY customer_id  LIMIT 6
")
```
### Exercise 6.3
Repeat the previous query and sort by the count in descending order.
```{r ex6.3}
dbGetQuery(con,"
 SELECT customer_id, COUNT(*) AS count
 FROM rental
 GROUP BY customer_id
 ORDER BY count DESC
 LIMIT 6
 ")
```

### Exercise 6.4
Repeat the previous query but use HAVING to only keep the groups with 40 or more.
```{r ex6.4}
dbGetQuery(con,"
 SELECT customer_id, COUNT(*) AS count
 FROM rental
 GROUP BY customer_id
  HAVING count > 40
  ORDER BY count DESC
")
```

## Exercise 7
The following query calculates a number of summary statistics for the payment table using MAX, MIN, AVG and SUM.
```{r ex7}
dbGetQuery(con,"
 SELECT MAX(amount) AS max,
        MIN(amount) as min,
        AVG(amount) AS avg,
        SUM(amount) AS sum
 FROM payment
")
```

### Exercise 7.1
Modify the above query to do those calculations for each customer_id.
```{r ex7.1}
dbGetQuery(con,"
 SELECT customer_id, 
        MIN(amount) as min,
        AVG(amount) AS avg,
        MAX(amount) AS max,
        SUM(amount) AS sum
 FROM payment
 GROUP BY customer_id
 LIMIT 10
")
```

### Exercise 7.2
Modify the above query to only keep the customer_ids that have more then 5 payments.
```{r ex7.2}
dbGetQuery(con,"
 SELECT customer_id, COUNT(*) AS count,
        MIN(amount) as min,
        AVG(amount) AS avg,
        MAX(amount) AS max,
        SUM(amount) AS sum
 FROM payment
 GROUP BY customer_id
 HAVING count  >   5
")
```

Cleanup:
Run the following chunk to disconnect from the connection.
 

```{r cleaning}
dbDisconnect(con)
```


Knit the document, commit your changes, and Save it on GitHub. 

git commit -a -m "Finalizing lab 10 https://github.com/USCbiostats/PM566/issues/47"


```{r sI}
sessionInfo()
```

