---
title: "03-lab-EDA"
author: "ks"
date: "9/10/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```

## Lab week 3


```{r install-libraries}
if (!require(data.table)) {install.packages("data.table")}
library(data.table)
library(tidyverse)
library(leaflet)
```

### 1. Read in the data

First read the data into a data.table.

```{r readdata}
if (!file.exists("met_all.gz")) {
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "data/met_all.gz", method="libcurl", timeout = 60)
}
met <- data.table::fread("met_all.gz")
```

### 2. Check the dimensions, header and footer.

```{r dim}
dim(met)
head(met)
tail(met)
```

How many rows and columns?

### 3. Look at variables

```{r str}
str(met)
```

### 4. Take a closer look

```{r table-year}
table(met$year)
```

```{r table-day}
table(met$day)
```

```{r temp}
summary(met$temp)
```

```{r elev}
summary(met$elev)
```

```{r wind}
summary(met$wind.sp)
```

#### Data cleaning

```{r correct-elev}
#met[met$elev==9999.0] <- NA
#met$elev[met$elev==9999.0] <- NA
met[met$elev==9999.0, elev:= NA ]
summary(met$elev)
```

At what elevation is the highest weather station?

The weather station with highest elevation is `r max(met$elev,na.rm=TRUE)` meters. This is after replacing 9999.0 values with the appropriate code for “missing”, which is “NA”.


Remove temps of -40 celsius.
```{r filter-temp}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```


### 5. Check the data against an external data source.

skip a few steps here...

### 6. Compute summary statistics

```{r elev-sum}
met[elev==max(elev,na.rm=TRUE)][, summary(wind.sp)]
met[elev==max(elev,na.rm=TRUE)][, summary(temp)]
```

```{r max-elev-site}
elev <- met[elev==max(elev,na.rm=TRUE)]
summary(elev)
```



```{r cor}
met[elev==max(elev,na.rm=TRUE), .(
  temp_wind = cor(temp,wind.sp,use="complete"),
  temp_day  =  cor(temp,day,use="complete"),
  wind_day  =  cor(wind.sp,day,use="complete")
)]
```

### 7. Exploratory graphs

We should look at the distributions of all of the key variables to make sure there are no remaining issues with the data.

```{r}
hist(met$elev, breaks=100)
```

```{r}
hist(met$wind.sp)
```

One thing we should consider for later analyses is to log transform wind speed and elevation as the are very skewed.

Look at where the weather station with highest elevation is located.

```{r}
#leaflet(elev) %>%
  #addProviderTiles('OpenStreetMap') %>% 
  #addCircles(lat=~lat,lng=~lon, opacity=1, #fillOpacity=1, radius=100)
```

The above doesn't render in .md.

Look at the time series of temperature and wind speed at this location. For this we will need to create a date-time variable for the x-axis.

```{r}
library(lubridate)
elev$date <- with(elev, ymd_hm(paste(year, month, day, hour, min, sep= ' ')))
summary(elev$date)
```

```{r}
elev <- elev[order(date)]
head(elev)
```

With the date-time variable we can plot the time series of temperature and wind speed.


```{r}
plot(elev$date, elev$temp, type='l')
#elev[,plot(date,temp,type="l")]
```

```{r}
plot(elev$date, elev$wind.sp, type='l')
```

Summarize any trends that you see in these time-series plots.

```{r}
elev2 <- elev[!is.na(wind.sp)]
plot(elev2$date, elev2$wind.sp, type='l')
lines(smooth.spline(elev2$date, elev2$wind.sp,
                    nknots=7),col=2,lwd=2)
```

###  SessionInfo

```{r}
sessionInfo()
```

