---
title: "03-lab-EDA"
author: "ks"
date: "9/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab week 3


```{r install-libraries}
if (!require(data.table)) {install.packages("data.table")}
library(data.table)
```

### 1. Read in the data

First read the data into a data.table.

```{r readdata}
if (!file.exists("met_all.gz")) {
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "data/met_all.gz", method="libcurl", timeout = 60)
}
met <- data.table::fread("met_all.gz")
```

### 2. Check the dimensions, header and footer.

```{r dim}
dim(met)
head(met)
tail(met)
```

How many rows and columns?

### 3. Look at variables

```{r str}
str(met)
```

### 4. Take a closer look

```{r table-year}
table(met$year)
```

```{r table-day}
table(met$day)
```

```{r temp}
summary(met$temp)
```

```{r elev}
summary(met$elev)
```

```{r wind}
summary(met$wind.sp)
```

#### Data cleaning

```{r correct-elev}
#met[met$elev==9999.0] <- NA
#met$elev[met$elev==9999.0] <- NA
met[met$elev==9999.0, elev:= NA ]
summary(met$elev)
```

At what elevation is the highest weather station?

The weather station with highest elevation is `r max(met$elev,na.rm=TRUE)` meters. This is after replacing 9999.0 values with the appropriate code for “missing”, which is “NA”.


Remove temps of -40 celsius.
```{r filter-temp}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```


### 5. Check the data against an external data source.

### 6. Compute summary statistics

```{r elev-sum}
met[elev==max(elev,na.rm=TRUE)][, summary(wind.sp)]
met[elev==max(elev,na.rm=TRUE)][, summary(temp)]
```

```{r max-elev-site}
elev <- met[elev==max(elev,na.rm=TRUE)]
summary(elev)
```



```{r cor}
met[elev==max(elev,na.rm=TRUE), .(
  temp_wind = cor(temp,wind.sp,use="complete"),
  temp_day  =  cor(temp,day,use="complete"),
  wind_day  =  cor(wind.sp,day,use="complete")
)]
```

### 7. Exploratory graphs

We should look at the distributions of all of the key variables to make sure there are no remaining issues with the data.

```{r}
hist(met$elev, breaks=100)
```




###  SessionInfo

```{r}
sessionInfo()
```

