---
title: "lab 6"
author: "ks"
date: "10/01/2021"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```

## Lab week 6 

```{r install-libraries}
library(data.table)
library(ggplot2)
library(dplyr)
library(tidytext)
library(tidyverse)
```


```{r get-data, cache=TRUE}
fn <- "mtsamples.csv"
if (!file.exists(fn))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv",
    destfile = fn)
mtsamples <- read_csv(fn)
#mtsamples <- as_tibble(fn)
```

Read in Medical Transcriptions
Loading in reference transcription samples from https://www.mtsamples.com/



### Question 1. What specialties do we have?

We can use count() from dplyr to figure out how many different categories do we have? Are these categories related? overlapping? evenly distributed?

```{r count, cache=TRUE}
specialties <- mtsamples %>%
   count(medical_specialty)
```

There are `r nrow(specialties)` specialties.

```{r table}
specialties %>%
  arrange(desc(n)) %>%
  top_n(15) %>%
  knitr::kable()
```

```{r dist1}
ggplot(mtsamples, aes(x = medical_specialty)) +
  geom_histogram(stat = "count") +
  coord_flip()
```

```{r dist2}
library(forcats)
ggplot(mtsamples, aes(n, fct_reorder(medical_specialties, n))) +
  geom_col()
```

These are not evenly (uniformly) distributed.

### Question 2. Tokenize the the words in the transcription column
Count the number of times each token appears
Visualize the top 20 most frequent words
Explain what we see from this result. Does it makes sense? What insights (if any) do we get?

```{r tokenize-transcription, cache=TRUE}
mtsamples %>%
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

The word "patient" seems to be important, but we observe a lot of stopwords.


### Question 3. 

Redo visualization but remove stop words before.
Bonus points if you remove numbers as well.
What do we see know that we have removed stop words? Does it give us a better idea of what the text is about?

```{r transcript-wo-stop-words, cache=TRUE}
mtsamples %>%
  unnest_tokens(output = word, input = transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = c("word")) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

Looking better, but we don't like the numbers.

 


### Question 4: 

repeat question 2, but this time tokenize into bi-grams. how does the result change if you look at tri-grams?

```{r bigrams}
#mtsamples %>%
#  unnest_ngrams(ngram, text, n=2)  %>%
#  count(ngram, sort = TRUE) %>%
#  top_n(20, n)  %>%
#  ggplot(aes(n, fct_reorder(ngram, n))) +
#  geom_col()
```

```{r trigrams}
#mtsamples %>%
#  unnest_ngrams(ngram, text, n=3)  %>%
#  count(ngram, sort = TRUE) %>%
#  top_n(20, n)  %>%
#  ggplot(aes(n, fct_reorder(ngram, n))) +
#  geom_col()
```

### Question 5.
Use the results you got from question 4. Pick a word and count the words that appears after and before it.

```{r count-before-words}
#mtsamples %>%
#  unnest_ngrams(ngram, text, n = 2) %>%
#  separate(ngram, into = c("word1", "word2"), sep = " ") %>%
#  select(word1, word2) %>%
##  filter(word2 == "alice") %>%
#  count(word1, sort = TRUE)
```

```{r count-after-words}
#mtsamples %>%
#  unnest_ngrams(ngram, text, n = 2) %>%
#  separate(ngram, into = c("word1", "word2"), sep = " ") %>%
#  select(word1, word2) %>%
#  filter(word1 == "alice") %>%
#  count(word2, sort = TRUE)
```

### Question 6.
Which words are most used in each of the specialties. you can use group_by() and top_n() from dplyr to have the calculations be done within each specialty. Remember to remove stopwords. How about the most 5 used words?

```{r q6}

```


Knit the document, commit your changes, and Save it on GitHub. Don’t forget to add README.md to the tree, the first time you render it.


Knit the doc and save it on GitHub.

git commit -a -m "Finalizing lab 5 https://github.com/USCbiostats/PM566/issues/23"


```{r sI}
sessionInfo()
```

